---
title: "Changes in Demographic after Armed Conflict"
author: "Alan Peral and Kadija Yilla"
date: "12/12/2017"
output:
  html_document: default
  pdf_document: default
bibliography: citations.bib
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, cache=TRUE, 
                      fig.width=7, fig.height=6, fig.align = "center")
options(digits=3)
```

```{r include=FALSE}
require(dplyr)
require(WDI)
require(plyr)
```

# Introduction

Since the end of the cold war in 1991, 2016 was the fifth most violent year in the world [@trends]. Currently in 2017, there are more than 40 active conflicts in the world right now. Armed conflicts severely affect the lives of individuals in those countries currently. Most of the data online, outlines the armed conflict and causlaities of the war. Others also include an [interactive feature](https://www.irinnews.org/maps-and-graphics/2017/04/04/updated-mapped-world-war) of the global conflict to observe the scale of the conflict, with respect to other nations. One [special interactive map](https://www.cfr.org/interactives/global-conflict-tracker#!/) illustrates how various armed conflict affect US interests. Despite all the information surrounding the conflict, none accurately depict how the conflict has affected the lives of the individuals alive and living amidst the war.

War impacts people profoundly, and yet countries like the US insist on engaging in numerous armed conflicts regardless of the outcome. Westernized people, on  account of the privilege awarded by not being engaged in the conflict on a daily basis, rarely underatand the impact of war on the population. How has the war impacted education? Are more males leaving school as a result of the conflict? What about women? How has the war impacted the need to have equal education for women? How has the conflict caused the governmental expenditures to changes? Is there more funding for the various industries or are a mojortiy of the funding being exploited to arm the conflict? In addition, do the citizens have access to water, electricity, and housing and basic human rights? 

[//]: # "insert picture of the shiny app that addresses one of these questions"

Our application, explored through the tabs, offers a tool to answer the questions as well as a variety of other questions. We intergrated the wealth of information available for armed conflicts and various findings on coutry-level data to provide a synthesized view of how armed conflicts affects various nations. By offering this perspective, we can observe that it severely affects countries in which the conflcit takes place and the other side is left unaffected and growing.

# Data

The first data set used is from [Kaggle](https://www.kaggle.com/sdorius/globses/data) titled "Country Socioeconomic Status Scores." This data set includes the overall score of socioeconomic sattus by country for every decade since 1880. Socioeconomic Status (SES) Scores measure the accesibility of an individual, household, or community to collective resources. This measure incoprorates, income, wealth, health, and occupation [@Psaki2014]. For each country, this indivudialzied measure is aggregated as a whole to provide the measure for the country.

```{r, echo=FALSE}
library(readr)
SES.data <- read_csv("GLOB.SES.csv")
SES.data %>% arrange(country, desc(year)) %>% select(c(2:6,8)) %>% head(5)
```

The second set is world data from the World Bank. The data was accessed from an r package `WDI`. Using the r package we choose 17 indicators which we believed were good measures of the well-being of the individuals in the country. The indicators were as follows:

* health expenditure, total (% of GDP)
* fertility rate, tota (births per total)
* life expectancy at birth, female (years)
* life expectancy at birth, total (years)
* mortality rate, under-5 (per 1000 live births)
* children in employment, total (% of children 7-14)
* labor force, female (% of total labor force)
* labor force participation rate, female (% of female population ages 15+)(modeled ILO estimate) indicator
* GINI index (World Bank estimate) indicator
* Refugee population by country or territory of origin ()
* Refugee population by country or territory of asylum ()
* Improved sanitation facilities (% of population with access) indicator
* Improved water source (% of population with access) indicator
* Access to electricity (% of population) indicator
* Population living in slums (% of urban population) indicator

```{r, echo=FALSE}
library(WDI)

# health expenditure, total (% of GDP) indicator = SH.XPD.TOTL.ZS
# fertility rate, tota (births per total) indicator = SP.DYN.TFRT.IN
# life expectancy at birth, female (years) indicator = SP.DYN.LE00.FE.IN
# life expectancy at birth, total (years) indicator = SP.DYN.LE00.IN
# mortality rate, under-5 (per 1000 live births) indicator = SH.DYN.MORT
# children in employment, total (% of children 7-14) indicator = SL.TLF.0714.ZS
# labor force, female (% of total labor force) indicator = SL.TLF.TOTL.FE.ZS
# labor force participation rate, female (% of female population ages 15+)(modeled ILO estimate) indicator = SL.TLF.CACT.FE.ZS
# GINI index (World Bank estimate) indicator = SI.POV.GINI
# Refugee population by country or territory of origin () indicator = SM.POP.REFG.OR
# Refugee population by country or territory of asylum () indicator = SM.POP.REFG
# Improved sanitation facilities (% of population with access) indicator = SH.STA.ACSN
# Improved water source (% of population with access) indicator = SH.H2O.SAFE.ZS
# Access to electricity (% of population) indicator = EG.ELC.ACCS.ZS
# Population living in slums (% of urban population) indicator = EN.POP.SLUM.UR.ZS


#stored the required variables in a vector to access all at once
ind = c("health_expenditure" = "SH.XPD.TOTL.ZS",
        "fetility_rate" = "SP.DYN.TFRT.IN",
        "life_expectancy_female" = "SP.DYN.LE00.FE.IN",
        "mortality_rate_under5" = "SH.DYN.MORT",
        "children_employment" = "SL.TLF.0714.ZS",
        "labor_force_female" = "SL.TLF.TOTL.FE.ZS",
        "labor_force_Particpation_rate" ='SL.TLF.CACT.FE.ZS',
        "gini_index" = "SI.POV.GINI",
        "refugee_origin" = "SM.POP.REFG.OR",
        "refugee_asylum" = "SM.POP.REFG",
        "sanitation_access" = "SH.STA.ACSN",
        "water_access" = "SH.H2O.SAFE.ZS",
        "electricity_access" = "EG.ELC.ACCS.ZS",
        "slums_population" = "EN.POP.SLUM.UR.ZS"
        )

WBD.data = WDI(indicator = ind, start=1960, end=2015)

#renamed variables to be more appropriate
rnm = c("SH.XPD.TOTL.ZS" = "health_expenditure",
        "SP.DYN.TFRT.IN" = "fetility_rate",
        "SP.DYN.LE00.FE.IN" = "life_expectancy_female",
        "SH.DYN.MORT" = "mortality_rate_under5",
        "SL.TLF.0714.ZS" = "children_employment",
        "SL.TLF.TOTL.FE.ZS" = "labor_force_female",
        "SL.TLF.CACT.FE.ZS" = "labor_force_Particpation_rate",
        "SI.POV.GINI" = "gini_index",
        "SM.POP.REFG.OR" = "refugee_origin",
        "SM.POP.REFG" = "refugee_asylum",
        "SH.STA.ACSN" = "sanitation_access",
        "SH.H2O.SAFE.ZS" = "water_access",
        "EG.ELC.ACCS.ZS" = "electricity_access",
        "EN.POP.SLUM.UR.ZS" = "slums_population"
        )

WBD.data2 <- rename(WBD.data, rnm)
WBD.data2 %>% select(2,3,5,6) %>% head(5)
```


The third and final set is data about armed conflicts from the Uppsala Conflict Data Program  (UCDP) at the department of Peace and Conflict Research, Uppsala University and the Centre for the Study of Civil War at the Peace Research Institute Oslo (PRIO). The  UCDP/PRIO Armed Conflict Dataset represents both internal and external conflict from 1946 to the present. It documents the two sides in the conflict, as well as the location of the conflict, and indicators of intensity and number of fatalities. Armed conflict is defined as [-@ConflictSet]:

> “A contested incompatibility that concerns government and/or
territory where the use of armed force between two parties, of which at least one is the
government of a state, results in at least 25 battle-related deaths in a calendar year" -- Gleditsch et al. (2002)

```{r, echo=FALSE}
library(readr)
conflict.data <- read_csv("ucdp-prio-acd-171.csv")
conflict.data %>% select(c(2,3,5,10)) %>% head(5)
```


### Wrangling

Using the three seperate sets, we were able to construct a comprehsnive data set of all necessary information using data wrangling techniques.First we merged the SES scores set, **SES.data**, with the WBD data, **WBD.data2** by merging based on the country first and then the year. Before the merge was successful, the country names had to be reconciled because of subtle differences in naming techniques. For example, "Korea, Dem. People’s Rep." in the WBD set had to be renamed to "North Korea" as that was the name in te SES set. We choose to rename the countries based on names we thought the general public would be more familiar with. 

```{r, echo=FALSE}
# we need to rename a couple of countries in both the SES dataset
# and the WB dataset. We normalize country names so that we can 
# merge the two datasets.

# renames in SES
# rename Lao to Laos 
SES.data$country[SES.data$country == "Lao"] <- "Laos"
# rename Congo, Dem Rep to Congo, Dem. Rep.
SES.data$country[SES.data$country == "Congo, Dem Rep"] <- "Congo, Dem. Rep."
# Trinidad & Tobago to Trinidad and Tobago
SES.data$country[SES.data$country == "Trinidad & Tobago"] <- "Trinidad and Tobago"

# renames in WBD

# rename Congo, Rep. to Congo
WBD.data2$country[WBD.data2$country == "Congo, Rep."] <- "Congo"
# rename Egypt, Arab Rep. to Egypt
WBD.data2$country[WBD.data2$country == "Egypt, Arab Rep."] <- "Egypt"
# rename Gambia, The to Gambia
WBD.data2$country[WBD.data2$country == "Gambia, The"] <- "Gambia"
# rename Hong Kong SAR, China to Hong Kong
WBD.data2$country[WBD.data2$country == "Hong Kong SAR, China"] <- "Hong Kong"
# rename Iran, Islamic Rep. to Iran
WBD.data2$country[WBD.data2$country == "Iran, Islamic Rep."] <- "Iran"
# rename Kyrgyz Republic to Kyrgyzstan
WBD.data2$country[WBD.data2$country == "Kyrgyz Republic"] <- "Kyrgyzstan"
# rename Korea, Rep. to South Korea
WBD.data2$country[WBD.data2$country == "Korea, Rep."] <- "South Korea"
# rename Lao PDR to Laos
WBD.data2$country[WBD.data2$country == "Lao PDR"] <- "Laos"
# rename Macao SAR, China to Macao
WBD.data2$country[WBD.data2$country == "Macao SAR, China"] <- "Macao"
# rename Russian Federation to Russia
WBD.data2$country[WBD.data2$country == "Russian Federation"] <- "Russia"
# rename Slovak Republic to Slovakia
WBD.data2$country[WBD.data2$country == "Slovak Republic"] <- "Slovakia"
# rename Syrian Arab Republic to Syria
WBD.data2$country[WBD.data2$country == "Syrian Arab Republic"] <- "Syria"
# rename Venezuela, RB to Venezuela
WBD.data2$country[WBD.data2$country == "Venezuela, RB"] <- "Venezuela"
# rename Yemen, Rep. to Yemen
WBD.data2$country[WBD.data2$country == "Yemen, Rep."] <- "Yemen"
# rename Korea, Dem. People's Rep to North Korea
WBD.data2$country[WBD.data2$country == "Korea, Dem. People’s Rep."] <- "North Korea"
# rename Korea, Rep. to South Korea
WBD.data2$country[WBD.data2$country == "Korea, Rep."] <- "South Korea"

#merging to the SES data set and the WBD data set
WBD.SES <- merge(WBD.data2, SES.data, by=c("country", "year"), all=TRUE)
WBD.SES %>% select(c(1,18:20)) %>% head(5)
```
Next we take the established **WBD.SES** set and merge with the **conflict.data** set based on the "country"/"location." In order to create a tidy data set upon merge, we had to clean-up the conflict set for the "location" variable. For some conflicts in which the conflict occured in different locations globally, we had to seperate the countries in the "location" field and recreate the data for the seperately valued contries. To achieve this we had to create a user-fucnction that replicated the speicified joined countries and duplicated the values in one step. For example, the India vs. Pakistan War of 1965 occured in both India and Pakistan but was represented as "India, Pakistan". 

```{r, echo=FALSE}
conflict.data %>% filter(location == "India, Pakistan") %>% select(c(2,3,5,10)) %>% head(5)
```
This was instead changed to:
```{r, echo=FALSE}
#function to duplicate multiple countries with the same values 
replicate_fcn_example <- function(location.vector) {
  conflict.expanded <- data.frame()
  split <- unlist(strsplit(location.vector, ", "))
  len_split <- length(split)   
  
  #find the row which has the stringed values 
  rownum <- unlist(which(conflict.data$location == location.vector))
  rowlength <- length(rownum)
  
  #replicate the specific row based on the row number 
  for(i in 1:rowlength){
    conflict.expanded <- rbind(conflict.expanded, 
                               conflict.data[rep(rownum[i],each=len_split),])
  }
  #sort by year
  conflict.expanded <- conflict.expanded[order(conflict.expanded$year),]  
  
  for(j in seq(1,nrow(conflict.expanded),1)){
    mod_val <- j %% len_split
    if ((mod_val) == 0) {
        mod_val <- len_split
      }
    conflict.expanded$location[j] <- split[mod_val]
  }
  conflict.expanded
}
```
```{r, echo=FALSE}
# at this point conflict.expanded will be a data frame
# with one set of expanded rows
# we want to now add this set of expanded rows to
# conflict.data and repeat the process for all the rows
# that need duplication
#WBD.data2 <- rename(WBD.data, rnm)
conflict.data <- read_csv("ucdp-prio-acd-171.csv")
WBD.SES <- merge(WBD.data2, SES.data, by=c("country", "year"), all=TRUE)

difference <- setdiff(conflict.data$location, WBD.SES$country)

#replicate the value as an example
conflict.expanded <- replicate_fcn_example(difference[6])
conflict.expanded %>% select(c(2,3,5,10)) %>% head(5)
```
In addition, we had to conduct a few name changes to reconcile the differences in set to have a bigger join. The final set which we will be working with is displayed below. The set excludes the regions created in the WBD set for summary purposed such as "Small States" and "Low Middle Income" countries. This set is comprehensive as it includes the SES work for the decade, other country-specific indicators, and information about the conflict for a specific year. 

```{r, echo=FALSE}
#function to duplicate multiple countries with the same values 
replicate_fcn <- function(location.vector) {
  conflict.expanded <- data.frame()
  split <- unlist(strsplit(location.vector, ", "))
  len_split <- length(split)   
  
  #find the row which has the stringed values 
  rownum <- unlist(which(conflict.data$location == location.vector))
  rowlength <- length(rownum)
  
  #replicate the specific row based on the row number 
  for(i in 1:rowlength){
    conflict.expanded <- rbind(conflict.expanded, 
                               conflict.data[rep(rownum[i],each=len_split),])
  }
  #sort by year
  conflict.expanded <- conflict.expanded[order(conflict.expanded$year),]  
  
  for(j in seq(1,nrow(conflict.expanded),1)){
    mod_val <- j %% len_split
    if ((mod_val) == 0) {
        mod_val <- len_split
      }
    conflict.expanded$location[j] <- split[mod_val]
  }
  
  #we now delete the original unreplicated rows from the dataset
  #numrow[1] has the first row number of the unreplicated row
  del.start <- rownum[1]
  #the last row
  del.end <- rownum[rowlength]
  del.list <- seq(del.start,del.end,1)
  #delete unreplicated rows
  conflict.data <- conflict.data[-del.list,]
  
  #append replicated rows to end of dataset
  conflict.data <- rbind(conflict.data, conflict.expanded)
  conflict.data
}

# at this point conflict.expanded will be a data frame
# with one set of expanded rows
# we want to now add this set of expanded rows to
# conflict.data and repeat the process for all the rows
# that need duplication

difference <- setdiff(conflict.data$location, WBD.SES$country)

#replicate all necessary values
conflict.data <- replicate_fcn(difference[3])
conflict.data <- replicate_fcn(difference[4])
conflict.data <- replicate_fcn(difference[6])
conflict.data <- replicate_fcn(difference[9])
conflict.data <- replicate_fcn(difference[10])
conflict.data <- replicate_fcn(difference[12])
conflict.data <- replicate_fcn(difference[13])
conflict.data <- replicate_fcn(difference[15])
conflict.data <- replicate_fcn(difference[17])
conflict.data <- replicate_fcn(difference[18])
conflict.data <- replicate_fcn(difference[19])
conflict.data <- replicate_fcn(difference[20])
conflict.data <- replicate_fcn(difference[22])
conflict.data <- replicate_fcn(difference[23])
conflict.data <- replicate_fcn(difference[25])
conflict.data <- replicate_fcn(difference[26])
conflict.data <- replicate_fcn(difference[27])
conflict.data <- replicate_fcn(difference[28])
conflict.data <- replicate_fcn(difference[30])
conflict.data <- replicate_fcn(difference[31])
conflict.data <- replicate_fcn(difference[32])
conflict.data <- replicate_fcn(difference[33])
conflict.data <- replicate_fcn(difference[34])
conflict.data <- replicate_fcn(difference[35])
conflict.data <- replicate_fcn(difference[36])
conflict.data <- replicate_fcn(difference[38])
conflict.data <- replicate_fcn(difference[39])
conflict.data <- replicate_fcn(difference[40])
conflict.data <- replicate_fcn(difference[41])
conflict.data <- replicate_fcn(difference[42])
conflict.data <- replicate_fcn(difference[43])
conflict.data <- replicate_fcn(difference[44])
conflict.data <- replicate_fcn(difference[45])
conflict.data <- replicate_fcn(difference[46])
conflict.data <- replicate_fcn(difference[47])
conflict.data <- replicate_fcn(difference[48])
conflict.data <- replicate_fcn(difference[49])
conflict.data <- replicate_fcn(difference[51])
conflict.data <- replicate_fcn(difference[54])
conflict.data <- replicate_fcn(difference[55])
conflict.data <- replicate_fcn(difference[56])
conflict.data <- replicate_fcn(difference[58])
conflict.data <- replicate_fcn(difference[59])
conflict.data <- replicate_fcn(difference[60])
conflict.data <- replicate_fcn(difference[61])
conflict.data <- replicate_fcn(difference[62])
conflict.data <- replicate_fcn(difference[63])
conflict.data <- replicate_fcn(difference[65])

#rename Cambodia (Kampuchea) to "Cambodia"
conflict.data$location[conflict.data$location == "Cambodia (Kampuchea)"] <- "Cambodia"
#rename "Russia (Soviet Union)" to "Russia"
conflict.data$location[conflict.data$location == "Russia (Soviet Union)"] <- "Russia"
#rename "Madagascar (Malagasy)" to "Madagascar"
conflict.data$location[conflict.data$location == "Madagascar (Malagasy)"] <- "Madagascar"
#rename North/South Yemen to Yemen
conflict.data$location[conflict.data$location == "Yemen (North Yemen)"] <- "Yemen"
conflict.data$location[conflict.data$location == "South Yemen"] <- "Yemen"
#rename North/South Vietnam to "Vietnam"
conflict.data$location[conflict.data$location == "Vietnam (North Vietnam)"] <- "Vietnam"
conflict.data$location[conflict.data$location == "South Vietnam"] <- "Vietnam"
#rename "United States of America" to "United States"
conflict.data$location[conflict.data$location == "United States of America"] <- "United States"
#rename "Myanmar (Burma)" to "Myanmar"
conflict.data$location[conflict.data$location == "Myanmar (Burma)"] <- "Myanmar"
#rename "DR Congo (Zaire)" to "Congo, Dem. Rep."
conflict.data$location[conflict.data$location == "DR Congo (Zaire)"] <- "Congo, Dem. Rep."
#rename "Rumania" to "Romania"
conflict.data$location[conflict.data$location == "Rumania"] <- "Romania"
#rename "Zimbabwe (Rhodesia)" to "Zimbabwe"
conflict.data$location[conflict.data$location == "Zimbabwe (Rhodesia)"] <- "Zimbabwe"
#rename "Serbia (Yugoslavia)" to "Serbia"
conflict.data$location[conflict.data$location == "Serbia (Yugoslavia)"] <- "Serbia"
#rename "Bosnia-Herzegovina" to "Bosnia and Herzegovina"
conflict.data$location[conflict.data$location == "Bosnia-Herzegovina"] <- "Bosnia and Herzegovina"
#rename "Ivory Coast" to "Cote d'Ivoire"
conflict.data$location[conflict.data$location == "Ivory Coast"] <- "Cote d'Ivoire"
#rename "Brunei" to "Brunei Darussalam"
conflict.data$location[conflict.data$location == "Brunei"] <- "Brunei Darussalam"

#merging to the WBD/SES data set and the conflict set
WBD.SES.conflict <- merge(conflict.data, WBD.SES, by.y=c("country", "year"), by.x=c("location", "year"), all=TRUE)


to.remove <- c("Early-demographic dividend", "East Asia & Pacific (IDA & IBRD countries)", "East Asia & Pacific (excluding high income)", "Europe & Central Asia (excluding high income)", "Europe & Central Asia (IDA & IBRD countries)", "Fragile and conflict affected situations", "Heavily indebted poor countries (HIPC)", "High income", "IBRD only", "IDA & IBRD total", "IDA blend", "IDA only", "IDA total", "Late-demographic dividend", "Latin America & Caribbean (excluding high income)", "Latin America & the Caribbean (IDA & IBRD countries)", "Least developed countries: UN classification", "Low & middle income", "Low income", "Lower middle income", "Middle East & North Africa (excluding high income)", "Middle East & North Africa (IDA & IBRD countries)", "Middle income", "Not classified", "OECD members", "Other small states", "Pacific island small states", "Post-demographic dividend", "Pre-demographic dividend", "Small States", "South Asia (IDA & IBRD)", "Sub-Saharan Africa (excluding high income)", "Sub-Saharan Africa (IDA & IBRD countries)", "Upper middle income")
WBD.SES.conflict <- WBD.SES.conflict[ !WBD.SES.conflict$location %in% to.remove,]

WBD.SES.conflict %>% select(c(1,2,4,35,45)) %>% filter(year == "1990") %>% head(5)
```



# Results

[//]: # "insert link to the shiny app"


### Limitations

issued with the conflict set
issues with "Taiwan" and "Hyderabad" in converting

# Conclusion

# References